A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-09 graph_syntax.tex
another round of edits

\section{Requirements for New Graph Syntax}
The ultimate goal of the new graph--based syntax presented here is to be able to 
fully describe the general structure of an MDAO problem indpendently of any solution information, 
while still being able to accomodate the more specific case when a solution 
strategy is applied. In order to achieve that goal 
the graph needs to accomodate a number of constructs of MDAO problems: 
\begin{itemize}
    \item Analysis tools and connections between them
    \item Design variables, objectives, and constraints
    \item Local and global properties
    \item Coupling between analyses
    \item Multi-fidelity analyses
\end{itemize}

Beyond those basic constructs, there are also three phases of a design process that 
all need to be representable with the new syntax. Firstly there is the initial problem definition
phase where the specific analysis tools and design goals are identified. At the end of this phase, 
a single formal problem formulation is selected specifying design variables, constraints, objectives, 
analysis tools, etc. Lastly some specifc procedure for solving the problem is selected, for example 
picking an MDAO optimization architecture. Using the proposed graph syntax, each of these phases 
can be represented with the following three graphs:
\begin{itemize}
    \item Maximal Connectivity Graph (MCG)
    \item Fundamental Problem  (FPG)
    \item Problem Solution Graph (SPG)
\end{itemize}

The \emph{maximal connectivity graph} represents the first phase with all 
analysis codes being considered and all possible connections between them also present. The second graph 
is the \emph{fundamental problem formulation graph}, which is the smallest possible graph 
that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
may be represented by including additional edges and nodes to represent the 
solution strategy being employed to solve the problem.

@@

\section{Syntax Definition}
In the rest of this section we present the necessary graph theory fundamentals to 
construct any of the different graphs discussed above. 
The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
For a \emph{directed graph*} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
of a set of sets. Each ordered pair represents an edge starting at the node 
indicated by the first entry and directed to the node indicated by the second 
entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=1.5in]{images/example_directed_graph}
	\end{center}
	\vspace{-20pt}
\caption{Example directed graph.}
\label{f:example directed graph}
\end{figure}
As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
\begin{IEEEeqnarray*}{rCl}
V & = & \{1,2,3,4,5\}, \\
E & = & \big\{(1,2),(3,2),(1,3),(3,4)\big\}.
\end{IEEEeqnarray*}

@@

In Alexandrov and Lewis's REMS syntax only two node types were present, variable 
and function, and only one edge type\cite{alexandrov2004}. We've chosen to rename the ``function'' node 
type to ``model'' to be more consistent with modern MDAO frameworks. This work 
adds two more node types and one more edge type to the syntax to allow descriptive
graphs for all three phases of the design process.

@@

\subsection{In-degree and Out-degree}
\label{s:indegree-outdegree}
The \emph{in-degree} of a node is the number of connection edges directed in and 
is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
The degree of a given node is only a function of the connection edges 
attached to it. The number of implicit edges is not relevant since any number 
of drivers could be involved, at different parts of an solution process. For 
example in a sequential optimization where a global optimizer is first applied
and a gradient based one is applied second, design variables would have incomming 
implicit edges from both optimizers but this would not affect the in-degree of those
variable nodes.

We also define the \emph{upper in-degree limit}

\begin{equation}
\txt{deg}_u^-(v):V \to \mathbb{N}
\end{equation} 
and the \emph{lower in-degree limit} as
\begin{equation}
\txt{deg}_l^-(v):V \to \mathbb{N}.
\end{equation}
These user-specified limits govern the number of connection edges that may be directed into a variable
node for a valid graph. Consider a variable node $v$ with
$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ must have exactly one
incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
these limit conditions: 
\begin{description}
  \item[hole: ] The number of incomming edges is less than the lower indegree limit:
    \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
\end{description} 
These violations will be managed in order to seek a working problem formulation.

\section{Graph Representation of MDAO Constructs}
%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
\subsection{Analysis Blocks and Connections}
Analysis tools take in a number input variables and then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by a group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. If constructing a weighted graph, the weight calculation edge 
provides the opportunity to computational cost. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed sub-graphs within a graph.

@@

Via the analysis block structure, we can also distinguish between input and ouptut
variable nodes, as shown in Fig.~\ref{f:analysis block}. Inputs are any variable nodes that have an outgoing edge into a model 
node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
edge joining them to a model node. These are labeled out as 
local inputs and local outputs. The variables are local because they are explicitely tied
to the analysis block and have no relevance without it.

MDAO problems require that information be passed between sets of analyses. When 
information from the output of one analysis block is passed, or connected, to the 
input of another analysis block, a new connection edge is added connecting the two 
variable nodes involved in the exchange. These connection edges are free, unlike the edges 
within an analysis block, they could be added or removed depending on the needs
of a given design problem. In other words, free connection edges are created by 
engineers linking the output of one tool to the input of another one. Note that 
it is allowed for a single output to have outgoing connection edges to multiple 
downstream inputs.

\subsection{Design Variables}
Design variables require external decisions 
to be made about a given system to identify them. Consider 
a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
which, could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
 $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
have no incomming edges and could each be selected as design variables. 
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
  \end{center}
  \caption{Notional graph with two potential design variables \label{f:designvars}}
\end{figure}

Given a MCG with the ${deg}_l^-(v)=1$ for all nodes, there will be a set of holes. 
Each hole represents a node that could potentially become a design variable. 
It is up to the designer to examine that variable and decide if it is appropriate to 
use it as a design variable. If a variable should be a design variable, then 
set it's ${deg}_l^-(v)=0$. However, some variable holes might not make good design 
variables. For instance, if your aircraft mission analysis code requires a drag input 
which ends up being a hole then it's likely that some new analysis tool is needed
to predict drag before the design process can continue. In that case, you leave
${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.

So by definition, a design variable is any variable node with ${deg}_l^-(v)=0$, 
When working with a MCG or FPG, only connection edges are allowed, so a design 
variable in either graph will be a variable with no edges. For a valid PSG 
any design variable node would have at least one, but possibly multiple, incomming 
implicit edges from one or more driver nodes.

\subsection{Objectives, and Constraints}
Objectives and constraints are also constructs that need to be identified by engineers 
as part of the design process. In the case of objective functions single output values 
could be selected, but commonly multiple values are assembeled together via simple 
algebraic expression to form a objective function. For example, the problem from 
Fig. \ref{f:designvars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
always given in the form of either an inequality of an equality, e.g. 
$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
algebraic expression that can be evaluated to produce an output.

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. So we define an expression block, as the collection of variable and model 
nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

@@

  A simple definition of a global node is any node that involves information 
  from more than one analysis block. Likewise, a local node is any node 
  that involves information from only a single analysis block. As stated before, variable 
  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
  and similarly model nodes are also local. Expression and driver nodes
  are not inherently local or global though. For these, their locality is determined 
  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.

@@

  Coupling exists in a design problem when a set of two or more analyses each depend on the
  others outputs. Consider the notional problem from section \ref{sec:spg_vs_fpg} where two 
  analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
  edge to $B.z$, indicating that there is a connection between the two. 
  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
  to be iterated on untill they converge.

@@

  It is important to note that the cycle in graph in Fig. \ref{f:coupling} does not contain 
  any driver nodes. That means that no solver, optimizer, or other iterative construct is 
  invovled in the coupling. This absence is intentional, as the coupling is an inherant part 
  of the problem. This corresponds directly To Eqn. \ref{eqn:simple_fpf}, where 
  each equality constraint is one edge in the graph. Transitionining to either 
  Eqn. \ref{eqn:simple} or Eqn. \ref{eqn:simple2} means selecting one of the nodes as the starting point 
  for the iteration, indicating the associated analysis block should be executed first.

  Larger problems can contain more complex cycles in thier FPG, indicating more 
  complex coupling between analyses. A cycle can contain more than 
  just two analysis blocks. Multiple independent cycles can exist, indicating 
  independent coupling relationships. Cycles can also overlap, where the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing all of this coupling 
  becomes difficult. From the perspective of the graph, the problem reduces down to 
  a ordering challenge. For each cycle in the graph, you need to pick a analysis block
  to start the iteration from. Rogers DEMAID tool uses a genetic algorithm to find an ordering 
  that minimizes the overall coupling of the system by separating independent 
  cycles in the graph \cite{rogers1996,rogers1996demaid}.

\subsection{Multi-fidelity Problems}

  A multi-fidelity problem can be characterized by two or more different analyses 
  each calculating the same data. Mapping this type of multi-fidelity situation 
  onto a graph will yeild two analysis blocks, each with thier own output 
  variable nodes, having outgoing edges to a third analysis block that needs the 
  data as input. Figure \ref{f:collision-example} shows a notional problem 
  that exhibits this structure, where the key feature is that variable node $y$ in 
  analysis block $C$ has two incomming edges.

@@

  When an input variable node has more than one incomming edge, designers must make
  some kind of decision about how to manage the flow of information in their problem. 
  In Eqn. \ref{e:collision} we defined a collision as the situation where
  there are more incomming connection edges than the ${deg}_u^-(y)$. So desginers can
  indicate how to handle conflicting edges in a problem graph by setting the ${deg}_u^-(y)$
  for a given node. In more concrete terms, setting this limit equates to defining 
  the part of the graph in question as single or multi-fidelity.

  When ${deg}_u^-(y)=1$, then any conflicting edges in the graph will cause a collision
  and a choice between the colliding analysis blocks must be made. This results in a 
  single fidelity design problem. On the other hand if ${deg}_u^-(y)=2$, then two 
  different analysis blocks can co-exist with a given graph without causing a 
  collision. The design problem is then a multi-fidelity problem.

@@

\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as
\begin{equation}
I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
\end{equation}
where the degree is calculated with respect to $F_2$. This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.

@@

%Let $F_2 = (V_M,C_{F,2})$ be an undirected graph, which is to say to disregard the directions associated with $C_{F,2}$.

%Fixed edges are fixed only with respect to the analysis block in which they reside, it is okay to remove them along with the entire analysis block.

>>>>

\section{Requirements for New Graph Syntax}
The ultimate goal of the new graph--based syntax presented here is to be able to 
fully describe the general structure of an MDAO problem independently of any solution information, 
while still being able to accomodate the more specific case when a solution 
strategy is applied. In order to achieve that goal 
the graph needs to accomodate a number of constructs of MDAO problems: 
\begin{itemize}
    \item Analysis tools and connections between them
    \item Design variables, objectives, and constraints
    \item Local and global properties
    \item Coupling between analyses
    \item Multi-fidelity analyses
\end{itemize}

Beyond those basic constructs, there are also three phases of a design process that 
all need to be representable with the new syntax. Firstly there is the initial problem definition
phase where the specific analysis tools and design goals are identified. At the end of this phase, 
a single formal problem formulation is selected specifying design variables, constraints, objectives, 
analysis tools, etc. Lastly some specific procedure for solving the problem is selected, for example 
picking an MDAO optimization architecture. Using the proposed graph syntax, each of these phases 
can be represented with the following three graphs:
\begin{itemize}
    \item Maximal Connectivity Graph (MCG)
    \item Fundamental Problem Graph (FPG)
    \item Problem Solution Graph (SPG)
\end{itemize}

The \emph{maximal connectivity graph} represents the first phase with all 
analysis codes being considered and all possible connections between them also present. The second graph 
is the \emph{fundamental problem graph}, which is the smallest possible graph 
that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
may be represented by including additional edges and nodes to represent the 
solution strategy being employed to solve the problem.

@@

\section{Syntax Definition}
In this section we present the necessary graph theory fundamentals to 
construct the graphs discussed above. 
The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
For a \emph{directed graph*} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
of a set of sets. Each ordered pair represents an edge starting at the node 
indicated by the first entry and directed to the node indicated by the second 
entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=1.5in]{images/example_directed_graph}
	\end{center}
	\vspace{-20pt}
\caption{Example directed graph.}
\label{f:example directed graph}
\end{figure}
As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
\begin{IEEEeqnarray*}{rCl}
V & = & \{1,2,3,4,5\}, \\
E & = & \big\{(1,2),(3,2),(1,3),(3,4)\big\}.
\end{IEEEeqnarray*}

@@

In Alexandrov and Lewis's REMS syntax only two node types were present, variable 
and function, and only one edge type\cite{alexandrov2004}. We've chosen to rename the ``function'' node 
type to ``model'' to be more consistent with modern MDAO frameworks. The present work 
adds two more node types and one more edge type to the syntax to allow descriptive
graphs for all three phases of the design process.

@@

\subsection{In-degree and Out-degree}
\label{s:indegree-outdegree}
The \emph{in--degree} of a node is the number of connection edges directed in and 
is denoted as $\txt{deg}^-(v)$, and the \emph{out--degree} 
is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
The degree of a given node is only a function of the connection edges 
attached to it. The number of implicit edges is not relevant since any number 
of drivers could be involved, at different parts of an solution process. For 
example in a sequential optimization where a global optimizer is first applied
and a gradient based one is applied second, design variables would have incomming 
implicit edges from both optimizers but this would not affect the in--degree of those
variable nodes. 
((DP:I started to remove any reference to the edge type in these definitions but then saw that you needed it. The reason I did that was because in--degree and out--degree are basic graph theory terms. What if the reference to implicit edges is handled with the upper in--degree limit, i.e. after the FPG is obtained the upper indegree limit for the design variable nodes is relaxed (increased) so that it isn't a problem what is or isn't counted?))

((DP:the definitions were in/out degree were not in my book, I had to get them from wikipedia, where they are spelled indegree and outdegree))

We also define the \emph{upper in--degree limit}

\begin{equation}
\txt{deg}_u^-(v):V \to \mathbb{N}
\end{equation} 
and the \emph{lower in--degree limit} as
\begin{equation}
\txt{deg}_l^-(v):V \to \mathbb{N}.
\end{equation}
These user-specified limits govern the number of connection edges that may be directed into a variable
node for a valid graph. Consider a variable node $v$ with
$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ must have exactly one
incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
these limit conditions: 
\begin{description}
  \item[hole: ] The number of incomming edges is less than the lower indegree limit:
    \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
	This represents a lack of information being supplied to the node.

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
	This represents an overload of information such that it is unclear which source to use.
\end{description} 
An algorithm for detecting and managing these violations if proposed in Sec.(()).

\section{Graph Representation of MDAO Constructs}
%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
\subsection{Analysis Blocks and Connections}
Analysis tools take in a number input variables and then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by a group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. This edge can be given a weight to represent the computational cost of the analysis block. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed sub-graphs within a graph.

@@

Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected. As shown in Fig.~\ref{f:analysis block}, inputs are any variable nodes that have an outgoing edge into a model 
node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
edge joining them to a model node. 
((DP:I'm not sure what the sentences before or after this comment mean))
These are labeled out as 
local inputs and local outputs. The variables are local because they are explicitely tied
to the analysis block and have no relevance without it.

MDAO problems require that information be passed between sets of analyses. When 
information from the output of one analysis block is passed, or connected, to the 
input of another analysis block, a new connection edge is added connecting the two 
variable nodes involved in the exchange. These connection edges are free---unlike the edges 
within an analysis block---which means they could be added or removed depending on the needs
of a given design problem. In other words, free connection edges are created by 
engineers linking the output of one tool to the input of another one. Note that 
it is allowed for a single output to have outgoing connection edges to multiple 
downstream inputs.

\subsection{Design Variables}
((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
Design variables require external decisions 
to be made about a given system to identify them. Consider 
a notional problem given by the graph in Fig.~\ref{f:designvars}. There are three input variables 
which could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
 $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
have no incomming edges and could each be selected as design variables. 
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
  \end{center}
  \caption{Notional graph with two potential design variables \label{f:designvars}}
\end{figure}

Given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, there will likely be a set of holes preventing a valid FPG from being obtained.
Each hole represents a node that could potentially become a design variable. 
It is up to the designer to examine each and decide if it is appropriate to 
consider it a design variable, in which case the designer shall set the lower indegree limit to be zero. However, some variable holes may not actually be suitable as a design variable. For instance, if an aircraft mission analysis code requires a drag input 
which ends up being a hole then it's likely best to leave
$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.

A design variable is defined is any variable node with $\txt{deg}_l^-(v)=0$. ((DP: it seems like this should be have been at the beginning of this subsection.)) 
When working with a MCG or FPG, connection edges are the only free edges, so a design 
variable in either graph will be a variable node ((only variable nodes that are local inputs)) with no incoming edges. For a valid PSG 
any design variable node would have at least one, but possibly multiple, incomming 
implicit edges from one or more driver nodes.

\subsection{Objectives, and Constraints}
Objectives and constraints are also MDAO constructs that need to be identified by engineers 
as part of the design process. In the case of objective functions single output values 
could be selected, but commonly multiple values are assembeled together via simple 
algebraic expression to form a objective function. For example, the problem depicted in 
Fig.~\ref{f:designvars} could have an objective function of $z^2$ or $2y + z$. Constraints are 
always given in the form of either an inequality of an equality, e.g. 
$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
algebraic expression that can be evaluated to produce an output.

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

@@

  A simple definition of a global node is any node that directly involves information 
  from more than one analysis block. Likewise, a local node is any node 
  that involves information from only a single analysis block. As stated before, variable 
  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
  and similarly model nodes are also local. Expression and driver nodes
  are not inherently local or global though. For these, their locality is determined 
  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.

@@

  Coupling exists in a design problem when a set of two or more analyses each depend on the
  others outputs. Consider the notional problem from Sec.~\ref{sec:spg_vs_fpg} where two 
  analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
  edge to $B.z$, indicating that there is a connection between the two. 
  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
  to be iterated on until they converge.

@@

  It is important to note that the cycle in graph in Fig.~\ref{f:coupling} does not contain 
  any driver nodes. That means that no solver, optimizer, or other iterative construct is 
  invovled in the coupling. This absence is intentional, as the coupling is an inherent part 
  of the problem. This corresponds directly to Eqn.~\ref{eqn:simple_fpf}, where 
  each equality constraint is one edge in the graph. Transitioning to either 
  Eqn.~\ref{eqn:simple} or Eqn.~\ref{eqn:simple2} means selecting one of the nodes as the starting point 
  for the iteration and indicating which associated analysis block should be executed first.

  Larger problems can contain more complex cycles in their FPG, indicating more 
  complex coupling between analyses. A cycle can contain more than 
  just two analysis blocks. Multiple independent cycles can exist, indicating 
  independent coupling relationships. Cycles can also overlap, where the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. From the perspective of the graph, the problem reduces down to 
  an ordering challenge. For each cycle in the graph, an analysis block needs to be selected 
  to start the iteration from. Rogers' DEMAID tool uses a genetic algorithm to find an ordering 
  that minimizes the overall coupling of the system by separating independent 
  cycles in the graph \cite{rogers1996,rogers1996demaid}. In the present work, Sec.(()) will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.

\subsection{Multi--fidelity Problems}

  A multi--fidelity problem can be characterized by two or more different analyses 
  each calculating the same data. Mapping this type of multi--fidelity situation 
  onto a graph will yield two analysis blocks, each with their own output 
  variable nodes, having outgoing edges to a third analysis block that needs the 
  data as input. Figure \ref{f:collision-example} shows a notional problem 
  that exhibits this structure, where the key feature is that variable node $y$ in 
  analysis block $C$ has two incoming edges.

@@

  When an input variable node has more than one incoming edge, designers must make
  some kind of decision about how to manage the flow of information in their problem. 
  In Eqn.~\ref{e:collision} we defined a collision as the situation where
  there are more incoming connection edges than the upper indegree limit. So desginers can
  indicate how to handle conflicting edges in a problem graph by setting the upper indegree limit
  for a given node. In more concrete terms, setting this limit equates to defining 
  the part of the graph in question as single or multi--fidelity.

  When $\txt{deg}_u^-(y)=1$, then any conflicting edges in the graph will cause a collision
  and a choice between the colliding analysis blocks must be made. This results in a 
  single fidelity design problem. On the other hand if $\txt{deg}_u^-(y)=2$, then two 
  different analysis blocks can co-exist with a given graph without causing a 
  collision. The design problem is then a multi-fidelity problem.

@@

\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as ((pruning))
\begin{equation}
I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
\end{equation}
where the degree is calculated with respect to $F_2$. This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.

@@
